Stack Safety is a Security Property

### 2021-04-16 tex/main.tex
Factored out "break condition"

@@

\definition
For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
let \(\components\) be the set of components \(\component\) such that
\(\domMap ~ \component = \outside\).
%That is, at the start of some call, we take every component in the stack.
A system enjoys \emph{stepwise stack confidentiality} if,
for all \(\nach\) and \(\NPCT\) such that \(\mach \approx_\components \nach\)
and \(\nach,\pol,\domMap \hookrightarrow \NPCT | \neg f_{|\rts|}\),
\(\MPCT \doteq \NPCT\) holds.

@@

In other words, stepwise stack confidentiality states that if we vary every component of the stack,
the resulting trace (starting in the same policy state) steps in lockstep with the
original trace until both have reached a point whose depth is below the starting
depth (this is a return).

@@

\definition
For any call trace \(\MPCT \in \MPCT^*\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\).
%
Then, supposing that \(\last(\MPCT) = \mach',\pol',\context'\),
let \(\components\) be the set of protected components that changed during
the call, \(\components = \mathit{protected}(\domMap,\rts) \cap \Delta(\mach, \mach')\),
and let \((\mach,\pol',\context') \hookrightarrow \MPCT'\).
A system enjoys {\em observational integrity} if for any \(\nach\) such that
\(\mach' \approx_{\components'} \nach\), and its induced trace
\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).

@@

Confidentiality is more sophisticated. We must capture the intuition that
secrets do not escape during a call {\em and} are not leaked during a return.
This require us to be able to quantify over components that preserve our \(\diamond\)
relation from earlier.

\definition Let \(\components\) be a set of components and \(\mach,\mach'\) and \(\nach,\nach'\)
be pairs of states. The {\em unsafe set} of \(\components\), written
\(\bar{\Diamond}(\components,\mach,\mach',\nach,\nach')\), is the set of all components
\(\component \in \Delta(\mach,\mach') \cup \Delta(\nach,\nach')\) where
\(\mach' ~ \component \not = \nach' ~ \component\).

\definition
For any call trace \(\MPCT \in \MPCT^*\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\) and
\(\components\) be the set of components \(\component\)
such that \(\domMap ~ \component \not = \outside\).
%
A system enjoys {\em observational confidentiality} if for all
\(\nach\) such that \(\mach \approx_\components \nach\) and
\(\nach \hookrightarrow \NPCT | \neg f_\depth \),
\(\MPCT\) and \(\NPCT\), the following conditions all hold:

\begin{itemize}
\item They are observationally equivalent, \(\obsof(\MPCT) = \obsof(\NPCT)\),
  meaning that no secrets leak during the call
\item \(f_\depth ~ \last(\MPCT) \leftrightarrow f_\depth ~
\last(\NPCT)\)---if one call returns, so does the other
\item Assuming both calls return, let \(\last(\MPCT) = \mach',\pol',\context'\)
  and \(\pi_\mach(\last(\NPCT)) = \nach'\).
  Let \(\components' = \bar{\Diamond}(\components,\mach,\mach',\nach,\nach'\).
  And let \(\nach''\) be a state such that \(\mach' \approx_\components \nach''\).
  Then if \(\mach',\pol',\context' \hookrightarrow \MPCT'\) and
  \(\nach',\pol',\context' \hookrightarrow \NPCT'\), it must be
  that \(\obsof(\MPCT') \simeq \obsof(\NPCT')\)
\end{itemize}

>>>>

\definition
For each trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \_,\_,(\_,\rts)\), the
{\em break condition} of \(\MPCT\) is a predicate on
states \(\mathit{brk}(\MPCT) = \neg f_{|\rts|}\). It follows from our definitions
that this holds on \(\MPCT\)'s final state, {\em unless it terminated prematurely due to a failstop}.

@@

\definition
Let the secret set of a context, \(\mathit{secret}(\domMap,\rts)\), be the
set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).

\definition
For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
let \(\components = \mathit{secret}(\context)\) and \(\nach\) be any variant
\(\mach \approx_\components \nach\).
A system enjoys \emph{stepwise stack confidentiality} if,
for \(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \mathit{brk}(\MPCT)\),
\(\MPCT \doteq \NPCT\) holds.

@@

In other words, stepwise stack confidentiality states that if we vary every component of the stack,
the resulting trace (starting in the same policy state) steps in lockstep with the
original trace until both have reached the break condition.

@@

\definition
For any call trace \(\MPCT \in \MPCT^*\),
let \(\head(\MPCT) = \mach,\pol,\context\).
%
Then, supposing that \(\last(\MPCT) = \mach',\pol',\context'\),
let \(\components\) be the set of protected components that changed during
the call, \(\components = \mathit{protected}(\context) \cap \Delta(\mach, \mach')\),
and let \((\mach,\pol',\context') \hookrightarrow \MPCT'\).
A system enjoys {\em observational integrity} if for any \(\nach\) such that
\(\mach' \approx_{\components'} \nach\), and its induced trace
\((\nach,\pol',\context') \hookrightarrow \NPCT\), \(\obsof(\MPCT') \simeq \obsof(\NPCT)\).

@@

Confidentiality is more sophisticated. We must capture the intuition that
secrets do not escape during a call {\em and} are not leaked during a return.
This require us to be able to quantify over components that fail to preserve
our \(\diamond\) relation.

\definition Let \(\mach,\mach'\) and \(\nach,\nach'\)
be pairs of states. Their {\em unsafe set}, written
\(\bar{\Diamond}(\mach,\mach',\nach,\nach')\), is the set of all components
\(\component \in \Delta(\mach,\mach') \cup \Delta(\nach,\nach')\) where
\(\mach' ~ \component \not = \nach' ~ \component\).

\definition
For any call trace \(\MPCT \in \MPCT^*\),
let \(\head(\MPCT) = \mach,\pol,\context\) and
\(\components = \mathit{secret}(\context)\).
%
A system enjoys {\em observational confidentiality} if for all
\(\nach\) such that \(\mach \approx_\components \nach\) and
\(\nach \hookrightarrow \NPCT | \mathit{brk}(\MPCT)\),
\(\MPCT\) and \(\NPCT\), the following conditions all hold:

\begin{itemize}
\item They are observationally equivalent, \(\obsof(\MPCT) = \obsof(\NPCT)\),
  meaning that no secrets leak during the call
\item \(\mathit{brk}(\MPCT) ~ \last(\MPCT) \leftrightarrow \mathit{brk}(\MPCT) ~
\last(\NPCT)\)---if one call returns, so does the other
\item Assuming both calls return, let \(\last(\MPCT) = \mach',\pol',\context'\)
  and \(\pi_\mach(\last(\NPCT)) = \nach'\).
  Let \(\components' = \components \cap \bar{\Diamond}(\mach,\mach',\nach,\nach')\).
  And let \(\nach''\) be a state such that \(\mach' \approx_\components \nach''\).
  Then if \(\mach',\pol',\context' \hookrightarrow \MPCT'\) and
  \(\nach',\pol',\context' \hookrightarrow \NPCT'\), it must be
  that \(\obsof(\MPCT') \simeq \obsof(\NPCT')\)
\end{itemize}
